\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2021-09-08 papers/icst22/rebuttal.md

For the final version, we propose to use the extra available space to clarify
the questions and concerns of the reviewers to the furthest extent. We will also
expand on implementation details and usage examples of MUTAGEN.

Reviewer #2
-----------

> given the absence of a direct comparison against FuzzChick, how do you make
> the case that your technique is an improvement over the state of the art?

We clarify that IFC stack machine case study shows a direct comparison between
FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of
mean-time-to-failure and failure rate. We will update the introduction to reflect
that explicitly.

@@

We argue that we are fair when comparing against FuzzChick for the following
reasons:

- Table 3 compares results only for the runs in which *FuzzChick actually finds
the injected bug* --- a distinction that benefits FuzzChick, since it shows a
low failure rate in this case study, and was only able to find 5 out of 20 bugs
with a 100% failure rate, whereas MUTAGEN achieves a 100% failure rate across all
bugs.

- Given the above observation, comparing only the MTTF across all runs
(bug-found or timeout) can skew the representation of the performance of both
tools since timed-out executions inflate the overall calculation. This is clear
in Fig. 4, where FuzzChick's MTTF clearly jumps over 10^5ms as soon as the
failure rate goes below 100%.

- We observed that FuzzChick tends to find bugs either relatively fast, or not
at all within the 1-hour testing budget. The unstable behavior of FuzzChick
could be alleviated by running it several times in a row using much shorter
timeout (e.g. 30 seconds). Under this consideration, for the comparison against
FuzzChick to be fair, our tool needs to be reliable (always find the bug)
without being (possibly orders of magnitude) slower on individual runs. Table 3
shows that this is the case, where the *A12* measure indicates that MUTAGEN is
not prone to need many more tests to find a bug that FuzzChick in most of the
cases (except only for bugs #15 and #18, where FuzzChick finds it more quickly
but not in all runs).

### 2021-09-22 papers/icst22/main.tex
Created folder for ICST submission
Fine pass over paper

%% These commands are for a JOURNAL article.
\acmConference[ICSE 2022]{The 44th International Conference on Software
  Engineering}{May 21–29, 2022}{Pittsburgh, PA, USA}

@@

\author{Agust\'in Mista}
\email{mista@chalmers.se}
\affiliation{%
  \institution{Chalmers University of Technology}
  \city{Gothenburg}
  \country{Sweden}
}

\author{Alejandro Russo}
\email{russo@chalmers.se}
\affiliation{%
  \institution{Chalmers University of Technology}
  \city{Gothenburg}
  \country{Sweden}
}

@@

In this work we present \mutagen, a CGPT framework that overcomes these
limitations by generating mutants in an \emph{exhaustive} manner.
%
This is coupled with heuristics that help to schedule mutants based on their
novelty and to minimize mutating subexpressions leading to scalability issues of
exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our results indicate that our approach is capable of outperforming existing CGPT
tools, as well as finding bugs in real-world scenarios.
\end{abstract}

@@

%% FuzzChick: PBT + code instrumentation + type-preserving mutations
%% \fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
%% for the Coq programming language that
%
Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
is a technique that borrows ideas from the fuzzing community to generate
highly-structured values while still using automatically derived generators.
%
%% Notably, this tool is implemented as an extension of \quickchick
%% \cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
%
Instead of continuously generating random invalid test cases from scratch, CGPT
works by keeping queues of \emph{interesting} previously executed test cases
that can be mutated using high-level, structure-preserving transformations to
produce new ones.
%
%% Mutations are achieved using high-level, structure-preserving transformations
%% that can be derived directly from the codebase.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a higher probability of producing a new interesting test
case than generating a new one using a na\"ive generator.
%
In addition, high-level, structure-preserving mutators are better suited for
producing syntactically valid mutants.
%
High-level mutators avoids wasting time with generic low-level mutators (e.g.,
at the bit level) that know very little about data structures, thus producing
syntactically broken mutants most of the time.
%
Such a grammar-aware mutation technique has shown to be quite useful when
fuzzing systems accepting structurally complex inputs \cite{holler2012fuzzing,
  wang2019superion, xsmith}.
%
While most of the tools following this itechnique use existing grammars to
tailor the mutators to the specific input structure used by the system under
test, external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
%% In CGPT, external grammars are not required.
%
The data types of the inputs to the testing properties already provide a good source
of information about how data can be mutated, e.g., by changing one data constructor
by another.
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
Specialized mutators can be automatically derived directly from data types
definitions --- making strongly-typed programming languages an ideal match for
CGPT.

@@

Worth mentioning, this technique is not meant to replace sophisticated, manually
written random generators, but rather to provide an adequate testing solution
for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%% Heuristics to improve testing loop
Additionally, the testing loop of \mutagen incorporates two heuristics that help
finding bugs faster and more reliably (\S \ref{sec:heuristics}).
%
In the first place, our tool uses first-in first-out (FIFO) scheduling with
priority to enqueue interesting test cases for mutation.
%
This scheduling algorithm is indexed by the novelty of each new test case with
respect to the ones already seen.
%
In this light, interesting test cases that discover new parts of the code
earlier during execution are given a higher priority.
%
Moreover, our scheduling lets the testing loop jump back and forth between
mutable test cases as soon as a new more interesting one is enqueued for
mutation, eliminating potential delays when the mutation queues grow large.

@@

In contrast to CGPT and many other popular fuzzing tools, \mutagen works by to
mutating test cases on an exhaustive and precise manner, where
%
\begin{inparaenum}
  \item each subexpression of a mutable test case is associated with a set of
    structure-preserving mutations, and
  \item each one of these mutations is scheduled \emph{exactly once}.
\end{inparaenum}
%
%% There is a small exception to this rules that we will introduce soon.
%
Our realization is that, by using an exhaustive mutation approach, we avoid
needing a heuristic to assign a mutation budget to each mutated test case.
%
Moreover, producing mutants exhaustively ensures that no interesting mutation is
omitted due to randomness, as well a no mutation is evaluated more than once.
%
This is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
\cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
produce test cases exhaustively --- please refer to \S \ref{sec:related}
for a detailed discussion.

@@

We evaluated the performance of \mutagen using two case studies.
%
The first one is a simple abstract stack machine that enforces
\emph{noninterference} \cite{goguen1982security} using runtime checks.
%
Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
systematically introducing 20 bugs on its enforcing mechanism.
%
\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
against RPBT using na\"ive- and hand-tuned random generators.
%
Here, we replicate their results and compare them against our tool.
%
Worth mentioning, we reimplemented this case study from scratch in Haskell in
order to run \mutagen on its test suite.

@@

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old a new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.
%
Later, bugs are systematically introduced in the IFC enforcing mechanism by
removing or weakening the checks stored in this rule table.

@@

In this work we use exhaustiveness as a way to reliably enforce that all
possible mutants of an interesting test case are executed.
%
In contrast to fully-exhaustive testing tools, \mutagen relies on randomness to
find initial interesting mutable test cases.
%
In terms of pruning, it is possible to instruct \mutagen to detect unused
subexpressions to avoid producing mutations over their corresponding positions.
%
This could improve the overall performance when testing non-strict properties
that leave parts of their input unevaluated.
%
In our case studies, however, we observed that properties we tested
%% their preconditions tend to be
are quite strict when executing test cases obtained by mutating valid existing ones,
fully evaluating their input before executing the postcondition.
%
Thus, we do not forsee considerable improvements by applying lazy-prunning in
this scenario --- gathering evidence about lazy pruning in \mutagen is as part
of our future work.

>>>>

%% These commands are for a JOURNAL article.
\acmConference[ICST 2022]{15th IEEE International Conference on Software
  Testing, Verification and Validation}{Apr 4–13, 2022}{}

@@

%% Hidden authors in the generated PDF.

%% \author{Agust\'in Mista}
%% \email{mista@chalmers.se}
%% \affiliation{%
%%   \institution{Chalmers University of Technology}
%%   \city{Gothenburg}
%%   \country{Sweden}
%% }

%% \author{Alejandro Russo}
%% \email{russo@chalmers.se}
%% \affiliation{%
%%   \institution{Chalmers University of Technology}
%%   \city{Gothenburg}
%%   \country{Sweden}
%% }

@@

In this work we present \mutagen, a CGPT framework that overcomes these
limitations by generating mutants in an \emph{exhaustive} manner.
%
This is coupled with heuristics that help to schedule mutants based on their
novelty and to minimize mutating certain subexpressions leading to scalability issues of
exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our results indicate that our approach is capable of outperforming existing CGPT
tools, as well as finding bugs in real-world scenarios.
\end{abstract}

@@

%% FuzzChick: PBT + code instrumentation + type-preserving mutations
%% \fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
%% for the Coq programming language that
%
Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
is a technique that borrows ideas from the fuzzing community to generate
highly-structured values while still using automatically derived generators.
%
%% Notably, this tool is implemented as an extension of \quickchick
%% \cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
%
Instead of continuously generating random invalid test cases from scratch, CGPT
works by keeping queues of \emph{interesting} previously executed test cases
that can be mutated using high-level, structure-preserving transformations to
produce new ones.
%
%% Mutations are achieved using high-level, structure-preserving transformations
%% that can be derived directly from the codebase.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a higher probability of producing a new interesting test
case than generating a new one using a na\"ive generator.
%
In addition, high-level, structure-preserving mutators are better suited for
producing syntactically valid mutants.
%
High-level mutators avoid wasting time with generic low-level mutators that work
at the bit level and know very little about data structures, thus producing
syntactically broken mutants most of the time.
%
Such a grammar-aware mutation technique has shown to be quite useful when
fuzzing systems accepting structurally complex inputs \cite{holler2012fuzzing,
  wang2019superion, xsmith}.
%
While most of the tools following this itechnique use existing grammars to
tailor the mutators to the specific input structure used by the system under
test, external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
%% In CGPT, external grammars are not required.
%
The data types of the inputs to the testing properties already provide a good source
of information about how data can be mutated, e.g., by changing one data constructor
by another.
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
Specialized mutators can be automatically derived directly from data types
definitions --- making strongly-typed programming languages an ideal match for
CGPT.

@@

Worth mentioning, this technique is not meant to replace sophisticated manually-written
random generators, but rather to provide an adequate testing solution
for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%% Heuristics to improve testing loop
Additionally, the testing loop of \mutagen incorporates two heuristics that help
finding bugs faster and more reliably (\S \ref{sec:heuristics}).
%
In the first place, our tool uses first-in first-out (FIFO) scheduling with
priority to enqueue interesting test cases for mutation.
%
This scheduling algorithm is indexed by the novelty of each new test case with
respect to the ones already seen.
%
In this light, interesting test cases that discover new parts of the code
earlier during execution are given a higher priority.
%
Moreover, our scheduling allows the testing loop to jump back and forth between
mutable test cases as soon as a new more interesting one is enqueued for
mutation, eliminating potential delays when the mutation queues grow large.

@@

In contrast to CGPT and many other popular fuzzing tools, \mutagen works by 
mutating test cases on an exhaustive and precise manner, where
%
\begin{inparaenum}
  \item each subexpression of a mutable test case is associated with a set of
    structure-preserving mutations, and
  \item each one of these mutations is scheduled \emph{exactly once}.
\end{inparaenum}
%
%% There is a small exception to this rules that we will introduce soon.
%
Our realization is that, by using an exhaustive mutation approach, we avoid
needing a heuristic to assign a mutation budget to each mutated test case.
%
Moreover, producing mutants exhaustively ensures that no interesting mutation is
omitted due to randomness, as well a no mutation is evaluated more than once.
%
This is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
\cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
produce test cases exhaustively --- please refer to \S \ref{sec:related}
for a detailed discussion.

@@

We evaluated the performance of \mutagen using two case studies.
%
The first one is a simple abstract stack machine that enforces
the hyperproperty \emph{noninterference} \cite{goguen1982security} using runtime checks.
%
Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
systematically introducing 20 bugs on its enforcing mechanism.
%
\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
against RPBT using na\"ive- and hand-tuned random generators.
%
Here, we replicate their results and compare them against our tool.
%
Worth mentioning, we reimplemented this case study from scratch in Haskell in
order to run \mutagen on its test suite.

@@

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old and new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.
%
Later, bugs are systematically introduced in the IFC enforcing mechanism by
removing or weakening the checks stored in this rule table.

@@

In this work we use exhaustiveness as a way to reliably enforce that all
possible mutants of an interesting test case are executed.
%
In contrast to fully-exhaustive testing tools, \mutagen relies on randomness to
find initial interesting mutable test cases.
%
In terms of pruning, it is possible to instruct \mutagen to detect unused
subexpressions to avoid producing mutations over their corresponding positions.
%
This could improve the overall performance when testing non-strict properties
that leave parts of their input unevaluated.
%
In our case studies, however, we observed that the properties we tested
%% their preconditions tend to be
are quite strict when executing test cases obtained by mutating valid existing ones,
fully evaluating their input before executing the postcondition.
%
Thus, we do not forsee considerable improvements by applying lazy-prunning in
this scenario --- gathering evidence about lazy pruning in \mutagen is as part
of our future work.
